---
title: "Post lab: Purrr Answer key"
author: "Nathaniel Grimes"
date: "2025-1-15"
format: 
  html:
    embed-resources: true
    code-fold: show
execute:
  warning: false
  eval: false
---

```{r}
library(tidyverse)
library(gapminder)
```

## Purrr Definitions

1. In your own words, describe what the map functions from the `purrr` package accomplish?

2. What data type do the following functions return?

  + `map()`
  + `map_chr()`
  + `map_dbl()`
  + `map_df()`
  + `map_lgl()`
  
  <span style="color:blue;">map() returns a list, map_chr() a character value, map_dbl() a double numeric value, map_df() a dataframe, and map_lgl() a logical results i.e. T/F</span>

3. If I have a dataset where I want to iterate over 4 variables that are changing, which of the following functions would I use?

  a. `map()`
  b. `map2()`
  c. `pmap()`
  d. `mutate()`

<span style="color:blue;">pmap() allows us to iterate over four different variables</span>

## Purrr Coding

1. Format the out of order code into the correct order for it to run

```{r}
output <- gapminder |>
  janitor::clean_names() |> 
  nest(.by=continent) |> 
  mutate(reg=map(.x=data, ~lm(life_exp~pop+gdp_per_cap+year,data=.x))) |> 
  mutate(coeff=map(.x=reg, ~coefficients(.x)))

```

2. Purr is excellent and speeding up simulation processes without having to rely on for loops. For example we could see how different production elasticities (a,b) benefit an economy with known capital (k) and labor (l). Run through the code below. Fix the error in `try1`. Which output would you prefer between a fixed `try1` and `try2`?


```{r}
parameters=expand_grid(a=seq(0.1,0.8,by=.1),b=seq(0.2,1,by=.1),k=4,l=2)

cobb<-function(a,b,k,l){
  out=k^a*l^b
  return(as.data.frame(out))
}

try1<-map2_df(.x=parameters$a,.y=parameters$b,~cobb(.x,.y,k=4,l=2))

try2<-pmap_df(parameters,cobb)

```


## Purrr in action

We want to plot 6 graphs for the top 6 most expensive child care states with data from 2008 to 2018. Each graph will be child care costs on the y-axis with two lines for median weekly infant and preschool care against the x axis of years. Modify the accompanying `plot_function` to be suitable for a `purrr` map. Create your own mapping and then use cowplot, ggarrange, or gridarrange to show all the plots on one figure. 

```{r}

# Run this code chunk to download data and get it in a good format
childcare_costs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/childcare_costs.csv') 

counties <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/counties.csv')

cc_clean<-childcare_costs |> 
  select(county_fips_code,mc_infant,mc_preschool,study_year) |> 
  full_join(counties)

top_6 <- cc_clean |>  
  group_by(state_name) |> 
  summarize(num=sum(is.na(mc_infant)),
            avg_cost=mean(mc_infant,na.rm=TRUE)) |> 
  filter(num==0) |> 
  top_n(6)

cc_plot_df <- cc_clean |> 
  filter(state_name %in% top_6$state_name) |> 
  group_by(state_name,study_year) |> 
  summarize(avg_infant=mean(mc_infant,na.rm=T),
            avg_pre=mean(mc_preschool,na.rm=T)) |> 
  pivot_longer(cols=c("avg_infant","avg_pre"),
               names_to = "childcare",
               values_to = "cost")
  
```



```{r pf_solution}
# Correct plot_function
plot_function<-function(df,name){
  p <- df |> 
    ggplot(aes(x=study_year,y=cost,color=childcare))+
      geom_line(linewidth=2)+
    scale_x_continuous(breaks=seq(2008,2018,by=2),labels = seq(2008,2018,by=2))+
    scale_color_manual(values=c("forestgreen","purple"),breaks=c("avg_infant","avg_pre"),labels=c("Infant","Preschool"),name="")+
      labs(x="",y="Average weekly\nchildcare cost",title =name)+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5))
  return(p)
}
```

```{r pf_results}

# How students should graph and use purrr
plot_list<-cc_plot_df |> 
  group_by(state_name) |> 
  nest() |> 
  mutate(plots=map2(.x=data,.y=state_name,~plot_function(.x,.y)))


ggpubr::ggarrange(plotlist=plot_list$plots, ncol=2, nrow=3, common.legend = TRUE, legend="bottom")


```

### Complete the following tasks with purrr

1. Generate 10 random samples from a normal distribution where the mean and standard deviation increase from 1 to 10 and  to 20. Hint: Make sure the sequences are of the same length by setting the length.out=11 argument in seq()

2. Use purrr to quickly return the class of every column from the `gapminder` dataset as characters

```{r}
# 1. Generate random draws

mean_in<-seq(1,10,length.out=11)
sd_in<-seq(2,20,length.out=11)

map2(mean_in,sd_in,~rnorm(n=10,mean=.x,sd=.y))

# 2. 

gapminder |> map_chr(class)

```

